### 2019-05-06 R/Analyse changes in pelagic stocks.Rmd
General commit of changes

# -----------------------------------------------------------------------------------------------
# Analyse changes in pelagic stocks.Rmd
#
# 07/01/2019 first coding; adapted from ICES advice and assessment summaries.Rmd
# 10/01/2018 made small changes to the table
# -----------------------------------------------------------------------------------------------

@@

  filter(grepl("her-47|mac-67|mac-nea|whb-c|whb-n|hom-c|hom-w|her-3a22|her-noss", 
               stockkeylabelold)) %>%
  filter(!grepl("benchmark", purpose)) %>% 
  filter(assessmentyear >= 2008) %>% 
  ungroup()

# create subset for assessments
e <-
  iAssess %>% 
  filter(stockkeylabelold %in% unique(d$stockkeylabelold)) %>% 
  filter(assessmentyear == 2018) %>% 
  filter(year >= 2008) %>% 
  filter(year <= assessmentyear) %>% 
  filter(adviceonstock == TRUE)

@@

d %>% 
  dplyr::select(stockkeylabelold, assessmentyear, fpa, fmsy) %>% 
  gather(key=var, value=value, fpa:fmsy) %>%

@@

d %>% 
  dplyr::select(stockkeylabelold, tacyear, adviceonstock, advisedlandingsmax) %>% 
  ggplot(aes(x=tacyear)) +
  theme_publication() +
  theme(legend.title = element_blank()) +

@@

  scale_x_continuous(breaks=pretty_breaks()) +
  expand_limits(y=0) + 
  labs(y="Advice Landings Max (tonnes)") +
  scale_color_manual(values = c("red", "blue")) +
  scale_linetype_manual(values = c("solid", "blank")) +
  facet_wrap(~stockkeylabelold, scales="free_y")

@@

d %>% 
  dplyr::select(stockkeylabelold, assessmentyear, adviceonstock, ssbay) %>% 
  rename(year = assessmentyear) %>%

@@

>>>>

# -----------------------------------------------------------------------------------------------
# Analyse changes in pelagic stocks.Rmd
#
# 07/01/2019 first coding; adapted from ICES advice and assessment summaries.Rmd
# 10/01/2018 made small changes to the table
# 04/05/2019 updated for WKRRMAC 2019; focus on mackerel
# -----------------------------------------------------------------------------------------------

@@

  filter(grepl("her-47|mac-67|mac-nea|whb-c|whb-n|hom-c|hom-w|her-3a22|her-noss", 
               stockkeylabelold)) %>%

  # count number of assessments per year; keep assessment if n=1 or otherwise if adviceonstock is TRUE
  group_by(stockkeylabelold, assessmentyear) %>% 
  mutate(n = n_distinct(paste0(stockkeylabelold, assessmentyear, adviceonstock))) %>% 
  # filter(n == 1 | ((n > 1 & adviceonstock == TRUE) | (n > 1 & benchmark == TRUE))) %>% 
  filter(n == 1 | (assessmentdate == max(assessmentdate, na.rm=TRUE)) ) %>% 
  # filter(!grepl("benchmark", purpose)) %>%

  filter(assessmentyear >= 2008) %>% 
  ungroup()

# d %>% filter(stockkeylabelold =="mac-nea", assessmentyear >= 2015) %>% View()

# create subset for assessments
e <-
  iAssess %>% 
  group_by(assessmentyear) %>% 
  filter(stockkeylabelold %in% unique(d$stockkeylabelold)) %>% 
  filter(assessmentyear == max(assessmentyear, na.rm=TRUE)) %>% 
  filter(year <= assessmentyear) %>%

  # count number of assessments per year; keep assessment if n=1 or otherwise if adviceonstock is TRUE
  group_by(stockkeylabelold, assessmentyear) %>% 
  mutate(n = n_distinct(paste0(stockkeylabelold, assessmentyear, adviceonstock))) %>% 
  filter(n == 1 | (assessmentdate == max(assessmentdate, na.rm=TRUE)) ) %>% 
  filter(year >= 2008) %>% 
  ungroup()
#   filter(n == 1 | (assessmentdate == max(assessmentdate, na.rm=TRUE)) )

# d %>%
#   ungroup() %>%
#   distinct(stockkeylabelold, assessmentyear,  benchmark, adviceonstock, n, .keep_all = FALSE) %>%
#   arrange(stockkeylabelold, assessmentyear, benchmark, adviceonstock, n) %>%
#   View()

@@

d %>% 
  dplyr::select(stockkeylabelold, assessmentyear, fpa, fmsy) %>% 
  gather(key=var, value=value, fpa:fmsy) %>% 
  # View()

@@

d %>% 
  dplyr::select(stockkeylabelold, tacyear, adviceonstock, advisedlandingsmax) %>%

  ggplot(aes(x=tacyear)) +
  theme_publication() +
  theme(legend.title = element_blank()) +

@@

  scale_x_continuous(breaks=pretty_breaks()) +
  expand_limits(y=0) + 
  labs(y="Advice Landings Max (tonnes)") +
  scale_color_manual(values = c("red", "blue")) +
  # scale_linetype_manual(values = c("solid", "blank", "dotted")) +
  facet_wrap(~stockkeylabelold, scales="free_y")

@@

d %>% 
  dplyr::select(stockkeylabelold, assessmentyear, adviceonstock, ssbay) %>% 
  rename(year = assessmentyear) %>% 
  View()

@@

# now based on assessment
t <-
  e %>% 
  filter(year == assessmentyear - 1 ) %>% 
  mutate(colour  = paste0(adviceonstock,benchmark),
         purpose = ifelse(grepl("bench", purpose), "benchmark", purpose),
         purpose = ifelse(grepl("initial", purpose), "replaced", purpose))

glimpse(d)

e %>% 
  group_by(stockkeylabelold) %>% 
  filter(assessmentyear == max(assessmentyear, na.rm=TRUE)) %>% 
  dplyr::select(stockkeylabelold, assessmentyear, year, adviceonstock, stocksize) %>% 
  # rename(year = assessmentyear) %>% 
  # View()

  # geom_point(aes(y=ssbay,
  #                colour   = ifelse(adviceonstock==TRUE,'advice','replaced'))) +
  geom_line(aes(y = stocksize )) +
  geom_point(aes(y = stocksize )) +

  geom_point(data=t, aes(x=year, y=stocksize, colour=purpose)) +
  scale_x_continuous(breaks=pretty_breaks()) +
  expand_limits(y=0) + 
  labs(y="SSB") +
  # scale_color_manual(values = c("red", "blue")) +
  facet_wrap(~stockkeylabelold, scales="free_y")
```

### 2019-05-06 R/paper 02 Scale-jump analysis of historical retrospective patterns.Rmd
General commit of changes

pairwise_comparisons <- function(removelastyears=0, numberofyears=10) {

@@

  res <- list("d.pcm"= d.pcm, 
              "d.pairwise.roughly"=d.pairwise.roughly,
              "d.pairwise" =d.pairwise)

@@

tmp <- pairwise_comparisons(removelastyears = 0, numberofyears=5)

@@

# plot scale jumps between assessments
res$d.pcm %>% 
  distinct(stock, variable, assessmentyear, phi) %>%
  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
  gather(key=metric, value=value, c(phi)) %>%

@@

  geom_line (aes(colour=factor(decade))) +
  geom_point(aes(colour=factor(decade), size = ifelse( abs(value) > threshold, 0.3, 0.0001))) +
  # geom_point(aes(colour=factor(decade), size = abs(value) )) +

@@

res$d.pcm %>% 
  distinct(stock, variable, assessmentyear, phi) %>%
  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
  gather(key=metric, value=value, c(phi)) %>%

@@

@@

@@

@@

d.jumps.byyear %>% 
  filter(assessmentyear >= 1990) %>%

@@

>>>>

pairwise_comparisons <- function(removelastyears=0, numberofyears=10, threshold=0.10) {

@@

    # calculate metrics
    filter(tmp=="firstyear") %>% 
    ungroup() %>% 
    select(stock, assessmentpair, variable, label, value) %>% 
    spread(key=label, value=value) %>% 
    mutate(phi            = (prior-last)/last,
           phi_abs        = abs(phi),
           assessmentyear = as.integer(substr(assessmentpair, 6,9)), 
           scale_jump     = ifelse(phi_abs  > threshold, TRUE, NA),
           scale_jump     = ifelse(phi_abs <= threshold, FALSE, scale_jump)) %>% 
    group_by(stock, variable) %>% 
    mutate(scale_jump     = ifelse(row_number() == 1, NA, scale_jump),
           removelastyears= removelastyears, 
           numberofyears  = numberofyears, 
           threshold      = threshold)

  res <- list("d.pcm"              = d.pcm, 
              "d.pairwise.roughly" = d.pairwise.roughly,
              "d.pairwise"         = d.pairwise)

@@

tmp <- pairwise_comparisons(removelastyears = 0, numberofyears=5, threshold = 0.1)

# tmp$d.pcm %>% View()

@@

res <- pairwise_comparisons(removelastyears = 0, numberofyears=10, threshold = 0.10)

# threshold <- 0.10

# plot scale jumps between assessments
res$d.pcm %>% 
  distinct(stock, variable, assessmentyear, phi, scale_jump, removelastyears, numberofyears, threshold) %>%
  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
  gather(key=metric, value=value, c(phi)) %>%

@@

  geom_line (aes(colour=factor(decade))) +
  geom_point(aes(colour=factor(decade), size = ifelse( scale_jump, 0.3, 0.0001))) +
  # geom_point(aes(colour=factor(decade), size = abs(value) )) +

@@

res$d.pcm %>% 
  distinct(stock, variable, assessmentyear, phi, scale_jump, removelastyears, numberofyears, threshold) %>%
  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
  gather(key=metric, value=value, c(phi)) %>%

@@

res <- pairwise_comparisons(removelastyears = 0, numberofyears=10, threshold=0.1)

@@

  geom_line (aes(colour=factor(decade))) +
  geom_point(aes(colour=factor(decade), size = ifelse( scale_jump, 0.3, 0.0001))) +

@@

# plot scale jumps with different settings for number of years and thresholds

d.pcm <- data.frame()

for (ny in c(5, 10, 15)) {
  for (nt in c(0.05, 0.1, 0.15, 0.2)) {
    res <- pairwise_comparisons(removelastyears = 0, numberofyears=ny, threshold=nt)
    d.pcm <- bind_rows(d.pcm, res$d.pcm)
  }
}

d.pcm %>% 
  filter(stock == "cod (north sea)") %>% 
  distinct(stock, variable, assessmentyear, phi, scale_jump, removelastyears, numberofyears, threshold) %>%
  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
  gather(key=metric, value=value, c(phi)) %>%

  ggplot(aes(x=assessmentyear, y=value, group=stock)) +
  theme_publication() +
  theme(
    panel.spacing    = unit(0.1, "lines"),
    panel.grid.major.x = element_line(colour = "gray80"),
    strip.background = element_blank(),
    strip.text.x     = element_text(face="bold", hjust=0.5, margin = margin(2,0,2,0, "mm")),
    strip.text.y     = element_text(face="bold", hjust=0.5, margin = margin(2,0,2,0, "mm"), angle = 0),
    strip.placement = "outside",
    legend.position  = "none"
  ) +

  geom_line (aes(colour=factor(decade))) +
  geom_point(aes(colour=factor(decade), size = ifelse( scale_jump, 0.3, 0.0001))) +

  labs(x=" ", y="") +
  facet_grid(threshold ~ numberofyears)

# plot number of scale jumps with different settings for number of years and thresholds

d.pcm <- data.frame()

for (ny in c(5, 10, 15)) {
  for (nt in c(0.05, 0.1, 0.15, 0.2)) {
    res <- pairwise_comparisons(removelastyears = 0, numberofyears=ny, threshold=nt)
    d.pcm <- bind_rows(d.pcm, res$d.pcm)
  }
}

d.pcm %>% 
  filter(stock == "cod (north sea)") %>% 
  distinct(stock, variable, assessmentyear, phi, scale_jump, removelastyears, numberofyears, threshold) %>%
  mutate(    decade = as.character(10*floor(assessmentyear/10)) ) %>% 
  gather(key=metric, value=value, c(phi)) %>%

  ggplot(aes(x=assessmentyear, y=value, group=stock)) +
  theme_publication() +
  theme(
    panel.spacing    = unit(0.1, "lines"),
    panel.grid.major.x = element_line(colour = "gray80"),
    strip.background = element_blank(),
    strip.text.x     = element_text(face="bold", hjust=0.5, margin = margin(2,0,2,0, "mm")),
    strip.text.y     = element_text(face="bold", hjust=0.5, margin = margin(2,0,2,0, "mm"), angle = 0),
    strip.placement = "outside",
    legend.position  = "none"
  ) +

  geom_line (aes(colour=factor(decade))) +
  geom_point(aes(colour=factor(decade), size = ifelse( scale_jump, 0.3, 0.0001))) +

  labs(x=" ", y="") +
  facet_grid(threshold ~ numberofyears)

@@

d.jumps.byyear %>% 
  filter(assessmentyear >= 1982) %>%

@@

d.assessments%>% 
  filter(assessmentyear <= 2018) %>% 
  group_by(assessmentyear) %>% 
  summarize(ncpueseries = mean(ncpueseries, na.rm=TRUE),
            nsurveyseries = mean(nsurveyseries, na.rm=TRUE)) %>% 
  gather(key=type, value=n, ncpueseries:nsurveyseries) %>%

  ggplot(aes(x=assessmentyear, y=n, group=type)) +
  theme_publication() +
  theme(
    panel.spacing    = unit(0.1, "lines"),
    panel.grid.major.x = element_line(colour = "gray80"),
    strip.background = element_blank(),
    strip.text.x     = element_text(size=8, face="bold", hjust=1, margin = margin(2,0,2,0, "mm")),
    strip.placement = "outside",
    legend.title    = element_blank()
  ) +
  # theme(legend.position  = "none") + 
  geom_line(aes(colour=type)) +
  labs(x="", y="avg")
